---
title: "Assignment4"
author: "ChenGuang,222017321102093"
date: "2020年5月18日"
output: html_document
---
***1.Build a classifier to predict labels `r` from `x` with xgboost, and show the confusion matrix***
Firstly, tidy the data. There are many NA in the data, we need to clean them.
```{r}
load("datasets/yrbs-1.rda")
na_position=which(is.na(r))
r1<-na.omit(r)
x1=x[-na_position,]

```
**Build a classifier**

Dividing the data into training set and test set.
```{r}
library(xgboost)
```

use cross validation to find best value of nrounds (and possibly eta)

```{r}
set.seed(2020-5-20)
xgb.cv(data =x1,label=r1,num_class=8,nrounds=20,nfold=10,objective="multi:softmax")
```
Add L2 regularisation penalty, prevent overfitting.
```{r}
xgb.cv(data =x1,label=r1,num_class=8,nrounds=20,nfold=10,objective="multi:softmax",lambda=1)
```
Seems to plateau at about 13 rounds. Similar performance after varying the learning rate and penalty
```{r}
model<-xgboost(data=x1,label = r1,num_class=8,nrounds = 13,objective="multi:softmax",lambda=1)
```
Confusion matrix:
```{r}
table(predict(model,newdata=x1),r1)
```
Calculate accuracy:9830/(5436+9830)=0.64391.
```{r}
table(predict(model,newdata=x1)==r1)
```

***2.Describe and visualise which variables are most important in the prediction. ***

**variable importance**
```{r}
xgb.importance(model = model)
xgb.plot.shap(model=model, data=x1,top_n=4,n_col=2)
```

For the table and these plots, we can see that **q97,q9,q99,q13** are most important variables.

***3.Describe and display the relationships between the most important variables and the label categories -- which category/categories is each of the most important variables useful for predicting? Can you produce a summary of the most distinctive predictors for each label category?***
```{r}
library(plyr)
library(tidyverse)
times<-function(x,times) {
    ifelse (x == 1, times[],
      ifelse (year(Date) == 2017, "2017","2018"))
}
ggplot(train,aes(label_train,train))+geom_bar(stat="identity",position="stack")
```

***4.Comment on whether (or not) task 3 would be ethically problematic if intended to be published, and for what reasons.***

(1) We can draw from the previous analysis that the accuracy of the prediction is not good, and it is easy to produce erroneous results.

(2) Deepen the stereotypes of different races, sometimes this is unreasonable.

(3) The prediction effect is not good, and the release will make people question the authenticity of the general survey and generate negative emotions for scientific analysis.

